Dealing with Consumer Uncertainty: Online Consumer Reviews and What makes them helpful?

### 2022-12-27 Paper.Rmd
little fixes after aprpoval

@@

## Negativity Bias Theorem and Loss Aversion

Negativity Bias Theorem argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. In @rozin2001's words Negativity Bias can be best exemplified such as the following:

@@

@rozin2001 explains the Negativity Bias Theorem by describing it with four main aspects. Namely, Negative Potency, Steeper Negative Gradients, Negative Dominance and Negative Differentiation [@rozin2001; @sen2007; @eslami2018]. Loss Aversion has a strong connection to Negativity Bias Theorem since negative cues might be used as an early warning against potential losses [@herr1991].

@@

On Figure \@ref(fig:Search-Neg) (A) and Figure \@ref(fig:Experience-Neg) (A) the distributions of the variable for both good types can be seen. On plots, X-axis labels were given in the logarithm of base 10 so 1.00 on the X-axis implies that 10% of the review was negative words. For example in this case the mean negative proportion rate is 2,3% ($10^{0.37}$) for search goods and 3,7% ($10^{0.57}$) for experience goods. It should be noted that `r round(mean(search_goods$negative == 0), digits = 2)` percent of the reviews for search goods do not have any negative words in them. For experience goods this number is `r round(mean(movies_data$negative == 0), digits = 2)` percent.

On Figure \@ref(fig:Search-Neg) (B), non-linear relationship between perceived helpfulness and the proportion of negative words can be observed for search goods. This might suggest that too much negative information might fall into slander rather than informing the customer about the potential risks of the product. On the other hand, somewhat weak negative linear relationship between perceived helpfulness and proportion of negative words can be seen for experience goods on Figure \@ref(fig:Experience-Neg) (B).

```{r Search-Neg, fig.cap="The Distribution(A) and the Scatter Plot(B) of Proportion of Negative Words for Search Goods"}
p1 <- search_goods %>% 
  ggplot(aes(negative))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(negative)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(search_goods$negative), y = Inf, 
           label = paste("Mean:", round(mean(search_goods$negative), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Count", x = "Log10(Proportion of Negative Words)")

p2 <- search_goods %>% 
  ggplot(aes(negative, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Helpfulness", x = "Log10(Proportion of Negative Words)")

@@

```{r Experience-Neg, fig.cap="The Distribution(A) and the Scatter Plot(B) of Proportion of Negative Words for Experience Goods"}
p1 <- movies_data %>% 
  ggplot(aes(negative))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(negative)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(movies_data$negative), y = Inf, 
           label = paste("Mean:", round(mean(movies_data$negative), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Count", x = "Log10(Proportion of Negative Words)")

p2 <- movies_data %>% 
  ggplot(aes(negative, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Helpfulness", x = "Log10(Proportion of Negative Words)")

@@

This variable was engineered in the same manner with its negative words counterpart. In this case, however, reviews of both good types have more positive words in them proportionally when compared to negative words. `r round(mean(search_goods$positive == 0), digits = 2)` percent of the reviews of search goods do not have any positive words in them and this number is only `r round(mean(movies_data$positive == 0), digits = 2)` for experience goods. Mean proportion values are 4,4% ($10^{0.64}$) and 5,2% ($10^{0.72}$) for search and experience goods respectively. Except for zero values both distributions are looking quite normal (see Figure \@ref(fig:Search-Pos) (A) and Figure \@ref(fig:Experience-Pos) (A)).

@@

On scatter plots (see Figure \@ref(fig:Search-Pos) (B) and Figure \@ref(fig:Experience-Pos) (B)), relationships between perceived helpfulness and proportion of positive words are looking very similar to their negative counterparts. Furthermore, these plots are kind of similar to perceived helpfulness versus number of words plots (see Figure \@ref(fig:Search-Length) (B) and Figure \@ref(fig:Experience-Length) (B)). In a sense, both negative and positive proportions imply the number of words. As the proportion increases so does the number of words. That is why aforementioned figures show similar relationships between variables. On higher dimensions these relationships might be different. That is why Exploratory Data Analysis should never be conducted in a confirmatory sense.

```{r Search-Pos, fig.cap="The Distribution(A) and the Scatter Plot(B) of Proportion of Positive Words for Search Goods"}
p1 <- search_goods %>% 
  ggplot(aes(positive))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(positive)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(search_goods$positive), y = Inf, 
           label = paste("Mean:", round(mean(search_goods$positive), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Count", x = "Log10(Proportion of Positive Words)")

p2 <- search_goods %>% 
  ggplot(aes(positive, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Helpfulness", x = "Log10(Proportion of Positive Words)")

@@

```{r Experience-Pos, fig.cap="The Distribution(A) and the Scatter Plot(B) of Proportion of Positive Words for Experience Goods"}
p1 <- movies_data %>% 
  ggplot(aes(positive))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(positive)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(movies_data$positive), y = Inf, 
           label = paste("Mean:", round(mean(movies_data$positive), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Count", x = "Log10(Proportion of Positive Words)")

p2 <- movies_data %>% 
  ggplot(aes(positive, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous(n.breaks = 8)+
  labs(y = "Helpfulness", x = "Log10(Proportion of Positive Words)")

@@

As it was mentioned in Section \@ref(negativity-bias-theorem-and-loss-aversion), people tend to try to minimize their losses. Because of this, negative cues draw more attention than positive ones since negative cues may imply potential losses. @guo2019 found out that people spent more time when evaluating negative reviews. This may suggest a linear interaction between Review Extremity and Review Length. Since people take more time to evaluate negative reviews, this may indicate that the effect of Review Length is stronger when Review Rating is low. And when a review is longer it is perceived as more helpful for negatively rated reviews than positively rated reviews. In this case, it is hypothesized that:

@@

Negativity Bias Theorem and Prospect Theory had a strong support in the case of search goods. The regression results showed that the reviews which had low ratings and/or reviews which had more negative words in them were found to be more helpful on the average. Additionally, Perceived Helpfulness has higher responsiveness to Review Length when a review has a low rating. This means that people focus more on negative cues than positive cues because the marginal value of an additional negative information is higher. This way consumers are able to cut down potential sure losses. A similar result was reflected by the negative coefficient of the proportion of positive words in a review. This shows that reviews with positive sentiments do not offer diagnostic value for consumers. Nonetheless, the U-Shape of the Review Rating variable also suggests positivity bias [@hoorens2014]. An explanation to this could be that certain brands or products might be favored by some customers and these customers might selectively tend towards positively rated reviews and find them helpful just because they love the brand [@zhu2021]. The reason why there is a divergence between the "positively rated review" and "a review with positive sentiment" might be explained by this. A consumer who has brand or product inclinations wants to be persuaded to have the product so this particular consumer does not really care about what is written in the review, he might skim through positive reviews until he feels persuaded. Since a preference towards a brand is subjective it might be speculated that a strong brand image could embed experience attributes into search goods. A similar notion was discussed by @mudambi2010 where they classified Apple iPod as an experience good. That is although an MP3 player such as iPod has many objective search attributes such as storage capacity and battery life, what makes this particular product preferable most of the time is the product image that was created by Apple.

@@

Online Consumer Reviews data were collected from two different sources (IMDB and BestBuy) and analyzed within the context of consumer uncertainty in this study. Negativity Bias Theorem, Prospect Theory and Search / Experience goods paradigm from information economics were used as theoretical engines to explain the information gathering process of an average consumer. Results showed similarities from the existing Review Helpfulness literature. However, once again it should be noted that the literature has many conflicting results.

This study contributes to the literature as an additional empirical work with a unique BestBuy review data and a unique combination of movie reviews from IMDB. In addition, to my knowledge, this is the first study which analyzed the linear continuous interaction between the Review Length and the Review Rating in terms of Negativity Bias Theorem. Additionally, there is a lack of emphasis on the sentiment analysis in the existing literature. By employing positive and negative sentiments in terms of Negativity Bias and Prospect Theory this study shows the viability of the sentiment analysis in this context.

>>>>

The existence of the Negativity Bias and how the potential effects of Negativity Bias on Review Helpfulness could be moderated by either Search And Experience goods will be the main focus of this study. A lesser degree of Negativity Bias is expected for Experience Goods. Additionally, the amount of information that a review has and its connection to Review Helpfulness with respect to Review Ratings will be analyzed. Next, the diagnostic values of the negative and positive sentiments are going to be discussed. Because of the Negativity Bias, a positive relationship between Negative Sentiments and Review Helpfulness is expected. Lastly, the relationship between the Review Helpfulness and the timeline of a review will be inquired.

@@

## Negativity Bias and Loss Aversion

Negativity Bias argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. In @rozin2001's words Negativity Bias can be best exemplified such as the following:

@@

@rozin2001 explains the Negativity Bias by describing it with four main aspects. Namely, Negative Potency, Steeper Negative Gradients, Negative Dominance and Negative Differentiation [@rozin2001; @sen2007; @eslami2018]. Loss Aversion has a strong connection to Negativity Bias since negative cues might be used as an early warning against potential losses [@herr1991].

@@

On Figure \@ref(fig:Search-Neg) (A) and Figure \@ref(fig:Experience-Neg) (A) the distributions of the variable for both good types can be seen. On logarithmic transformed plots, X-axis labels were given in the logarithm of base 10 so 1.00 on the X-axis implies that 10% of the review was negative words. For example in this case the mean negative proportion rate is 2,3% ($10^{0.37}$) for search goods and 3,7% ($10^{0.57}$) for experience goods. It should be noted that `r round(mean(search_goods$negative == 0), digits = 2)` percent of the reviews for search goods do not have any negative words in them. For experience goods this number is `r round(mean(movies_data$negative == 0), digits = 2)` percent.

On Figure \@ref(fig:Search-Neg) (D), non-linear relationship between perceived helpfulness and the logarithm of the proportion of negative words can be observed for search goods. This might suggest that too much negative information might fall into slander rather than informing the customer about the potential risks of the product. On the other hand, somewhat weak negative linear relationship between perceived helpfulness and proportion of negative words can be seen for experience goods on Figure \@ref(fig:Experience-Neg) (D).

```{r Search-Neg, fig.height=6, fig.cap="The Distribution(A) and the Scatter Plot(C) of Proportion of Negative Words for Search Goods and their respective logaritmic transformations (B), (D)"}
p1 <- search_goods %>% 
  ggplot(aes(negative))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(negative)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(search_goods$negative), y = Inf, 
           label = paste("Mean:", round(mean(search_goods$negative), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Log10(Proportion of Negative Words)")

p3 <- search_goods %>% 
  mutate(negative = 10^(negative)) %>% 
  ggplot(aes(negative))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(negative)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = 10^mean(search_goods$negative), y = Inf, 
           label = paste("Mean:", round(10^mean(search_goods$negative), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Proportion of Negative Words")

p2 <- search_goods %>% 
  ggplot(aes(negative, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  labs(y = "Helpfulness", x = "Log10(Proportion of Negative Words)")

p4 <- search_goods %>% 
  mutate(negative = 10^negative) %>% 
  ggplot(aes(negative, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  coord_cartesian(ylim = c(-0.1,1))+
  labs(y = "Helpfulness", x = "Proportion of Negative Words")

plots <- (p3 + p1) / (p4 + p2)

@@

```{r Experience-Neg, fig.height=6, fig.cap="The Distribution(A) and the Scatter Plot(C) of Proportion of Negative Words for Experience Goods and their respective logarithmic transformations (B), (D)"}
p1 <- movies_data %>% 
  ggplot(aes(negative))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(negative)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(movies_data$negative), y = Inf, 
           label = paste("Mean:", round(mean(movies_data$negative), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Log10(Proportion of Negative Words)")

p3 <- movies_data %>% 
  mutate(negative = 10^negative) %>% 
  ggplot(aes(negative))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(negative)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = 10^mean(movies_data$negative), y = Inf, 
           label = paste("Mean:", round(10^mean(movies_data$negative), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Proportion of Negative Words")

p2 <- movies_data %>% 
  ggplot(aes(negative, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  labs(y = "Helpfulness", x = "Log10(Proportion of Negative Words)")

p4 <- movies_data %>% 
  mutate(negative = 10^negative) %>% 
  ggplot(aes(negative, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  labs(y = "Helpfulness", x = "Proportion of Negative Words")

plots <- (p3 + p1) / (p4 + p2)

@@

This variable was engineered in the same manner with its negative words counterpart. In this case, however, reviews of both good types have more positive words in them proportionally when compared to negative words. `r round(mean(search_goods$positive == 0), digits = 2)` percent of the reviews of search goods do not have any positive words in them and this number is only `r round(mean(movies_data$positive == 0), digits = 2)` for experience goods. Mean proportion values are 4,4% ($10^{0.64}$) and 5,2% ($10^{0.72}$) for search and experience goods respectively. Except for zero values both distributions are looking quite normal (see Figure \@ref(fig:Search-Pos) (B) and Figure \@ref(fig:Experience-Pos) (B)).

@@

On scatter plots (see Figure \@ref(fig:Search-Pos) (D) and Figure \@ref(fig:Experience-Pos) (D)), relationships between perceived helpfulness and proportion of positive words are looking very similar to their negative counterparts. Furthermore, these plots are kind of similar to perceived helpfulness versus number of words plots (see Figure \@ref(fig:Search-Length) (B) and Figure \@ref(fig:Experience-Length) (B)). In a sense, both negative and positive proportions imply the number of words. As the proportion increases so does the number of words. That is why aforementioned figures show similar relationships between variables. On higher dimensions these relationships might be different. That is why Exploratory Data Analysis should never be conducted in a confirmatory sense.

```{r Search-Pos, fig.height=6, fig.cap="The Distribution(A) and the Scatter Plot(C) of Proportion of Positive Words for Search Goods and their respective logarithmic tansformations(B), (D)"}
p1 <- search_goods %>% 
  ggplot(aes(positive))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(positive)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(search_goods$positive), y = Inf, 
           label = paste("Mean:", round(mean(search_goods$positive), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Log10(Proportion of Positive Words)")

p3 <- search_goods %>% 
  mutate(positive = 10^positive) %>% 
  ggplot(aes(positive))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(positive)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = 10^mean(search_goods$positive), y = Inf, 
           label = paste("Mean:", round(10^mean(search_goods$positive), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Proportion of Positive Words")

p2 <- search_goods %>% 
  ggplot(aes(positive, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  labs(y = "Helpfulness", x = "Log10(Proportion of Positive Words)")

p4 <- search_goods %>% 
  mutate(positive = 10^positive) %>% 
  ggplot(aes(positive, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  coord_cartesian(ylim = c(0,1))+
  labs(y = "Helpfulness", x = "Proportion of Positive Words")

plots <- (p3 + p1) / (p4 + p2)

@@

```{r Experience-Pos, fig.height=6, fig.cap="The Distribution(A) and the Scatter Plot(C) of Proportion of Positive Words for Experience Goods and their respective logarithmic transformations (B),(D)"}
p1 <- movies_data %>% 
  ggplot(aes(positive))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(positive)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = mean(movies_data$positive), y = Inf, 
           label = paste("Mean:", round(mean(movies_data$positive), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Log10(Proportion of Positive Words)")

p3 <- movies_data %>% 
  mutate(positive = 10^positive) %>% 
  ggplot(aes(positive))+
  geom_histogram(color = "white", fill = "steelblue", boundary = 0)+
  geom_vline(aes(xintercept = mean(positive)), size = 1, alpha = 0.6)+
  annotate(geom = "label", x = 10^mean(movies_data$positive), y = Inf, 
           label = paste("Mean:", round(10^mean(movies_data$positive), digits = 2)),
           vjust = 1, hjust = 0, size = 4.5)+
  scale_x_continuous()+
  labs(y = "Count", x = "Proportion of Positive Words")

p2 <- movies_data %>% 
  ggplot(aes(positive, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  labs(y = "Helpfulness", x = "Log10(Proportion of Positive Words)")

p4 <- movies_data %>% 
  mutate(positive = 10^positive) %>% 
  ggplot(aes(positive, Helpfulness))+
  geom_point(alpha = 0.1)+
  geom_smooth(color = "steelblue", size = 1.7)+
  scale_x_continuous()+
  labs(y = "Helpfulness", x = "Proportion of Positive Words")

plots <- (p3 + p1) / (p4 + p2)

@@

As it was mentioned in Section \@ref(negativity-bias-and-loss-aversion), people tend to try to minimize their losses. Because of this, negative cues draw more attention than positive ones since negative cues may imply potential losses. @guo2019 found out that people spent more time when evaluating negative reviews. This may suggest a linear interaction between Review Extremity and Review Length. Since people take more time to evaluate negative reviews, this may indicate that the effect of Review Length is stronger when Review Rating is low. And when a review is longer it is perceived as more helpful for negatively rated reviews than positively rated reviews. In this case, it is hypothesized that:

@@

Negativity Bias and Prospect Theory had a strong support in the case of search goods. The regression results showed that the reviews which had low ratings and/or reviews which had more negative words in them were found to be more helpful on the average. Additionally, Perceived Helpfulness has higher responsiveness to Review Length when a review has a low rating. This means that people focus more on negative cues than positive cues because the marginal value of an additional negative information is higher. This way consumers are able to cut down potential sure losses. A similar result was reflected by the negative coefficient of the proportion of positive words in a review. This shows that reviews with positive sentiments do not offer diagnostic value for consumers. Nonetheless, the U-Shape of the Review Rating variable also suggests positivity bias [@hoorens2014]. An explanation to this could be that certain brands or products might be favored by some customers and these customers might selectively tend towards positively rated reviews and find them helpful just because they love the brand [@zhu2021]. The reason why there is a divergence between the "positively rated review" and "a review with positive sentiment" might be explained by this. A consumer who has brand or product inclinations wants to be persuaded to have the product so this particular consumer does not really care about what is written in the review, he might skim through positive reviews until he feels persuaded. Since a preference towards a brand is subjective it might be speculated that a strong brand image could embed experience attributes into search goods. A similar notion was discussed by @mudambi2010 where they classified Apple iPod as an experience good. That is although an MP3 player such as iPod has many objective search attributes such as storage capacity and battery life, what makes this particular product preferable most of the time is the product image that was created by Apple.

@@

Online Consumer Reviews data were collected from two different sources (IMDB and BestBuy) and analyzed within the context of consumer uncertainty in this study. Negativity Bias, Prospect Theory and Search / Experience goods paradigm from information economics were used as theoretical engines to explain the information gathering process of an average consumer. Results showed similarities from the existing Review Helpfulness literature. However, once again it should be noted that the literature has many conflicting results.

This study contributes to the literature as an additional empirical work with a unique BestBuy review data and a unique combination of movie reviews from IMDB. In addition, to my knowledge, this is the first study which analyzed the linear continuous interaction between the Review Length and the Review Rating in terms of Negativity Bias. Additionally, there is a lack of emphasis on the sentiment analysis in the existing literature. By employing positive and negative sentiments in terms of Negativity Bias and Prospect Theory this study shows the viability of the sentiment analysis in this context.
